D-optimal Designs for Logistic Regression using a Genetic Algorithm

### 2021-06-18 paper/paper.tex
finished introduction

There are many papers on optimal designs for logistic regression. The locally D-optiaml design for the lienar predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically. Ford (1972) proposed a D-optimal design on an unbounded interval \textbf{CHECH}. Sebastinani and Settimi (1992) derive D-optimal designs on several bounded design intervals. Numerical appproaches have also been treid. For example, Lall et. al (2018) find D-optimal designs on using a modified Fedorov algorithm.

For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results. Fornious (2005) found a D-optimal design for the quadratic model numerically.

Although some optimal designs for simple models have been obtained analytically, the optimal designs for more complicated models may be difficult or impossible to obtain in this way. Traditional numerical optimization algorithms may help, they may also be complicated to implement. In both apporaches, minor changes to design may have major consequences in the ability to obtain the opitmal design.

>>>>

There are many papers on optimal designs for logistic regression. The locally D-optiaml design for the lienar predictor $\eta = \beta_0 + \beta_1 x$ can be found analytically. Ford (1972) proposed a D-optimal design on an unbounded interval \textbf{CHECH}. Sebastinani and Settimi (1992) derive D-optimal designs on several bounded design intervals. Numerical appproaches have also been tried. For example, Lall et. al (2018) find D-optimal designs on using a modified Fedorov algorithm.

For the quadratic model with linear predictor $\eta = \beta_0 + \beta_1 + \beta_2x^2$, there have been no analytic results. Fornious (2005) found locally D-optimal designs numerically for the quadratic model on an unbounded interval. Results for a bounded interval are unknown.

For more complicated models, such as the quadratic logistic model, it can ber difficult an optimal design analytically. Numerical approaches such as by Fornious also rely on significant analytical work to find the optimal value of the objective function. Small changes to the model or to the design interval might result in a completely new problem with new analytical challenges.

Another approach is to use a general purpose optimization algorithm to find the optimal designs. However, some algorithms are better suited for the task. For example, it can difficult to take the gradient of the objective function for many optimal design problems. This is more difficult by the fact that changing the model will result in a new gradient that has to be derived. Therefore, it is better to use a gradient-free algorithm for these problems. Another problem that arises is local optima. Thus, we will want to use an optimization algorithm that can ``explore" the optimization space without getting stuck in local optima.

Nature-inspired metaheuristics are a family of algorithms that are gradient free and have mechanisms for escaping local optima. Examples include particle swarm optimization (Kennedy, Eberhart,  1995), the artificial bee colony algorithm (Karaboğa, 2005), and differential evolution (Storn, Price, 1997). These algorithms mimic the behavior of animals or nature processes to find a global optimum.

In this paper, we use the genetic algorithm (Holland, 1975) which is inspired by natural selection. The algorithm starts with an initial population of solution vectors and then combines and mutates the solutions to hopefully produce more solutions that have a better objective function value. The genetic algorithm has many different implementations and is widely applied to solve optimization problems. In this paper, we use the genetic algorithm to find locally D-optimal designs for several logistic regression models.
